---
title: "scallops - trawl data set"
author: "Mike Allen"
date: "10/29/2021"
output: html_document
---

# load libraries and data
```{r}
library(tidyverse)
library(sf)
library(rnaturalearth)
library(here)
is.odd <- function(x) x %% 2 != 0

# read in trawl data, filter to just scallops (sent by Fred K)
dat <- readRDS("data_noaa/trawl/dat.rds") %>%
  filter(grepl(sppocean, pattern = "placopecten"))

# read in haul data (sent by Fred K)
# subset hauls to only include NEFSC_NEUS which has length
# other surveys with scallops:
# "DFO_ScotianShelf", "VIMS_NEAMAP", "DFO_Newfoundland" 
# "SCDNR_SEUS" is embedded within, but has no scallops
hauls_df_step1 <- readRDS("data_noaa/trawl/hauls.rds") %>%
  filter(region %in% c("NEFSC_NEUS")) %>%
  # fill all haul data with scallop wtcpue
  left_join(dat, by = "haulid") %>%
  select(-sppocean, -Freq, -presfit, -logwtcpue) %>%
  replace_na(list(wtcpue = 0)) %>%
  mutate(decade = case_when(year<1990 ~ "1980s", 
                            year>=1990 & year <2000 ~ "1990s",
                            year>=2000 ~ "2000s-2010s"),
         grid_lat = case_when(lat-trunc(lat)<0.5 ~ trunc(lat)+0.25,
                              lat-trunc(lat)>=0.5 ~ 
                                trunc(lat)+0.75),
         grid_lon = case_when(abs(lon-trunc(lon))<0.5 ~ 
                                trunc(lon)-0.25,
                              abs(lon-trunc(lon))>=0.5 ~ 
                                trunc(lon)-0.75),
         grid = paste0(grid_lat,grid_lon)) %>%
  arrange(haulid, year)

# Read in NEUS survey length data from OceanAdapt 
# originally obtained from Alexa F as "neus_Survdat.RData"
tlen <- readRDS("data_noaa/trawl/survdat.rds") %>%
  # filter to just scallops (401)
  filter(SVSPP == 401) %>%
  # divide into length bins (measures coarser than dredge survey)
  mutate(
    n_0_70 = ifelse(LENGTH <= 6, NUMLEN, 0),
    n_70_100 = ifelse(LENGTH > 6 & LENGTH < 10, NUMLEN, 0),
    n_100_up = ifelse(LENGTH >= 10, NUMLEN, 0),
    station_temp = case_when(nchar(STATION)==1 ~ 
                               paste0("00", as.character(STATION)),
                             nchar(STATION)==2 ~
                               paste0("0", as.character(STATION)),
                             nchar(STATION)==3 ~
                               as.character(STATION),
                             TRUE ~ "problem"),
    haulid = paste(CRUISE6,station_temp,STRATUM, sep = "-")) %>%
  select(-station_temp)

# include total catch number from tlen in hauls_df
hauls_df <- tlen %>%
  group_by(haulid) %>%
  summarize(n_catch_all = sum(NUMLEN, na.rm = T), # 23 NA, 1963: 17
            n_0_70 = sum(n_0_70),
            n_70_100 = sum(n_70_100), 
            n_100_up = sum(n_100_up), 
            .groups = "drop") %>%
  right_join(hauls_df_step1, by = "haulid") %>%
  replace_na(replace = list(n_catch_all = 0,
                            n_0_70 = 0,
                            n_70_100 = 0,
                            n_100_up = 0)) %>%
  arrange(haulid, year)
# note: haulid "201404-211-1100" has EXTREMELY high count (>150k)

# # make a data frame of all length-haulid combos for DRM age model
# tlen$length = as.numeric(floor(tlen$LENGTH))
# len = unique(tlen$length)[order(unique(tlen$length))][2:21]
# 
# hauls_drm_list <- list()
# for(i in len){
#   print(i)
# hauls_drm_list[[i]] <- hauls_df_step1 %>%
#   mutate(length = i,
#          spp = "Placopecten magellanicus") %>%
#   select(haulid, length, spp, btemp = bottemp, 
#          year, month, lat, lon) %>%
#   left_join(select(tlen,
#                    haulid,
#                    length,
#                    number_at_length = NUMLEN),
#                    by = c("haulid", "length")) %>%
#   select(haulid, length, spp, number_at_length,
#          btemp, year, month, lat, lon) %>%
#   replace_na(., list(number_at_length=0))
# }
# 
# hauls_drm <- do.call(rbind, hauls_drm_list)
# write.csv(hauls_drm, "output/scallop_catch_at_length_mo.csv",
#           row.names = F)
hauls_drm <- read.csv("output/scallop_catch_at_length_mo.csv")

# rm(hauls_df_step1)

# summarizing hauls data by 0.5 degree grid cell
hd_df = hauls_df %>%
    mutate(lat = grid_lat,
         lon = grid_lon) %>%
    group_by(grid, year, decade, lat, lon, 
           grid_lat, grid_lon) %>%
  summarize(mean_wtcpue = mean(wtcpue),
            mean_tot_num = mean(n_catch_all),
            mean_0_70 = mean(n_0_70),
            mean_70_100 = mean(n_70_100),
            mean_100_up = mean(n_100_up),
            n_hauls = length(wtcpue),
            mean_depth = mean(elev, na.rm = T),
            mean_bottemp = mean(bottemp, na.rm = T),
            .groups = "drop") %>%
  arrange(grid, year) %>%
  filter(lat >34)

# make data frames into sf objects and load spatial data
hauls <- hauls_df %>%
  st_as_sf(coords = c("lon", "lat"),
           crs = 4326)

hd <- hd_df %>%
  st_as_sf(coords = c("grid_lon", "grid_lat"),
           crs = 4326)

state_bounds <- ne_coastline(scale = 10, returnclass = "sf") %>%
  st_crop(hd)

rm(hauls_df, dat)

# summarize change in mean catch between 1980-1999 and 2000-2015
change_plot = hd %>%
  st_drop_geometry() %>%
  as.data.frame() %>%
  filter(year >= 1980) %>%
  mutate(decade = case_when(decade == "2000s-2010s" ~ 
                              "Years_2000_2015",
                            TRUE ~ "Years_1980_1999")) %>%
  group_by(grid, lat, lon, decade) %>%
  summarize(mean_catch = mean(mean_wtcpue, na.rm = T),
            .groups = "drop") %>%
  pivot_wider(id_cols = c(grid, lat, lon), names_from = decade,
              values_from = "mean_catch") %>%
  mutate(
    pct_change = 
           (100*Years_2000_2015 - Years_1980_1999)/Years_1980_1999,
    change = Years_2000_2015 - Years_1980_1999,
    grid_lat = lat, grid_lon = lon
    ) %>%
  st_as_sf(coords = c("grid_lon", "grid_lat"),
           crs = 4326) 

# Read and format climate data
clim1 <- readxl::read_xlsx(here("data_noaa/climate/",
                               "Scallops grid statistics.xlsx"),
                          skip = 1) %>%
  pivot_longer(cols = 3:142) %>%
  rename(grid_lat = Latitude, grid_lon = Longitude) %>%
  mutate(grid = paste0(grid_lat, grid_lon))

clim <- clim1 %>%
  mutate(year = rep(c(sapply(1980:2014, 
                             function(x) rep(x, 4))), 
                    length(unique(clim1$grid))),
         name = case_when(grepl(name, pattern = "Min t") ~ 
                                "min_temp",
                              grepl(name, pattern = "Max t") ~ 
                                      "max_temp",
                              grepl(name, pattern = "Avg t") ~ 
                                      "mean_temp",
                              grepl(name, pattern = "O2") ~ 
                                      "O2"
                              )) %>%
  pivot_wider(id_cols = c(grid, grid_lon, grid_lat, year))

rm(clim1)
# write.csv(clim, "data_noaa/climate/climate_formatted.csv",
#           row.names = F)

# change in climate between 2 periods
change_clim <- clim %>% 
  mutate(decade = case_when(year >= 2000 ~ 
                              "Years_2000_2015",
                            TRUE ~ "Years_1980_1999")) %>%
  group_by(grid, grid_lat, grid_lon, decade) %>%
  summarize(mean_bottemp = mean(mean_temp, na.rm = T),
            .groups = "drop") %>%
  pivot_wider(id_cols = c(grid, grid_lat, grid_lon), 
              names_from = decade,
              values_from = "mean_bottemp") %>%
  mutate(
    change = Years_2000_2015 - Years_1980_1999,
    pct_change = change/Years_1980_1999,
    lat = grid_lat, lon = grid_lon
    ) %>%
  st_as_sf(coords = c("grid_lon", "grid_lat"),
           crs = 4326)

# change in min temp (new data) between 2 periods
change_clim_min <- clim %>% 
  mutate(decade = case_when(year >= 2000 ~ 
                              "Years_2000_2015",
                            TRUE ~ "Years_1980_1999")) %>%
  group_by(grid, grid_lat, grid_lon, decade) %>%
  summarize(min_bottemp = mean(min_temp, na.rm = T),
            .groups = "drop") %>%
  pivot_wider(id_cols = c(grid, grid_lat, grid_lon), 
              names_from = decade,
              values_from = "min_bottemp") %>%
  mutate(
    change = Years_2000_2015 - Years_1980_1999,
    pct_change = change/Years_1980_1999,
    lat = grid_lat, lon = grid_lon
    ) %>%
  st_as_sf(coords = c("grid_lon", "grid_lat"),
           crs = 4326)

change_O2 <- clim %>% 
  mutate(decade = case_when(year >= 2000 ~ 
                              "Years_2000_2015",
                            TRUE ~ "Years_1980_1999")) %>%
  group_by(grid, grid_lat, grid_lon, decade) %>%
  summarize(mean_O2 = mean(O2, na.rm = T),
            .groups = "drop") %>%
  pivot_wider(id_cols = c(grid, grid_lat, grid_lon), 
              names_from = decade,
              values_from = "mean_O2") %>%
  mutate(
    change = Years_2000_2015 - Years_1980_1999,
    pct_change = change/Years_1980_1999,
    lat = grid_lat, lon = grid_lon
    ) %>%
  st_as_sf(coords = c("grid_lon", "grid_lat"),
           crs = 4326)

# combine Morely climate data with new climate data to compare
clim.both <- hauls_df_step1 %>%
  group_by(grid, year) %>%
  summarise(SBT.actual = mean(SBT.actual),
            SBT.max = mean(SBT.max),
            SBT.min = mean(SBT.min)) %>%
  right_join(clim, by = c("grid", "year"))

# Moreley data: change in climate between 2 periods
change_morely <- clim.both %>% 
  mutate(decade = case_when(year >= 2000 ~ 
                              "Years_2000_2015",
                            TRUE ~ "Years_1980_1999")) %>%
  group_by(grid, grid_lat, grid_lon, decade) %>%
  summarize(mean_bottemp = mean(SBT.actual, na.rm = T),
            .groups = "drop") %>%
  pivot_wider(id_cols = c(grid, grid_lat, grid_lon), 
              names_from = decade,
              values_from = "mean_bottemp") %>%
  mutate(
    change = Years_2000_2015 - Years_1980_1999,
    pct_change = change/Years_1980_1999,
    lat = grid_lat, lon = grid_lon
    ) %>%
  st_as_sf(coords = c("grid_lon", "grid_lat"),
           crs = 4326)
```
# Plot 0.5 x 0.5 degree grid cell averages - mean catch by decade
```{r}
# summarize mean catch by decade
hd %>%
  group_by(grid, lat, lon, decade) %>%
  summarize(mean_catch = mean(mean_wtcpue),
            .groups = "drop") %>%
ggplot() +
  geom_tile(aes(x = lon, y = lat, fill = log10(mean_catch+1)), 
            alpha = 1, size = 0.25) +
  geom_sf(data = state_bounds, aes(), color = "gray",
          size = .25) +  theme_bw() +
  theme(text = element_text(size = 15),
        axis.text = element_text(size = 6),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8)) +
  viridis::scale_fill_viridis(option = "inferno") +
  labs(x = "", y = "", fill = "log10(wt+1)") +
  facet_wrap(~decade)

# ggsave("figures/scallop_trawl_hd_mean_catch_by_decade.png",
#         height = 3, width = 7, dpi = 400)
```
# Plot CHANGE in average catch by time period (0.5 grid cell)
"Because of these measures, the abundance of sea scallops
recovered rapidly and was fully rebuilt by 2001 (Hart &
Rago, 2006)." Zang 2022
```{r}
# summarize change in mean catch between 1980-1999 and 2000-2015
change_plot = hd %>%
  st_drop_geometry() %>%
  as.data.frame() %>%
  filter(year >= 1980) %>%
  mutate(decade = case_when(decade == "2000s-2010s" ~ 
                              "Years_2000_2015",
                            TRUE ~ "Years_1980_1999")) %>%
  group_by(grid, lat, lon, decade) %>%
  summarize(mean_catch = mean(mean_wtcpue, na.rm = T),
            .groups = "drop") %>%
  pivot_wider(id_cols = c(grid, lat, lon), names_from = decade,
              values_from = "mean_catch") %>%
  mutate(
    pct_change = 
           (100*Years_2000_2015 - Years_1980_1999)/Years_1980_1999,
    change = Years_2000_2015 - Years_1980_1999,
    grid_lat = lat, grid_lon = lon
    ) %>%
  st_as_sf(coords = c("grid_lon", "grid_lat"),
           crs = 4326) 

ggplot() +
  geom_point(aes(x = lon, y = lat, fill = change,
                 size = Years_2000_2015), 
            # alpha = 1,
            data = change_plot,
            shape = 21,
            color = "black") +
  geom_point(aes(x = lon, y = lat,
                 size = Years_1980_1999), 
            # alpha = 1,
            data = change_plot,
            shape = 1,
            color = "black") +
  geom_sf(data = state_bounds, aes(), color = "gray",
          size = .25) +  theme_bw() +
  theme(text = element_text(size = 15),
        axis.text = element_text(size = 6),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8)) +
  viridis::scale_color_viridis(option = "inferno") +
  labs(x = "", y = "", fill = "Absolute\nchange",
       size = "Mean CPUE\nin each\nperiod",
       title = "Change in sea scallop catch,\ntrawl survey (1980-1999 vs. 2000-2015)")+ 
  scale_fill_gradient2(midpoint = 0, 
                       low = "steelblue", 
                       mid = "white",
                       high = "firebrick", space = "Lab") +
  theme(text = element_text(size = 14),
        title = element_text(size = 12))

# ggsave("figures/scallop_trawl_hd_change_in_mean_catch.png",
#         height = 5, width = 5, dpi = 400)
```
# Plot AVERAGE catch by time period (0.5 grid cell)
```{r}
# plot mean abundance, 2000-2015
(abund_per2 <- ggplot() +
  geom_tile(aes(x = lon, y = lat, fill = Years_2000_2015),
            # alpha = 1,
            data = change_plot,
            color = "black") +
  geom_sf(data = state_bounds, aes(), color = "gray",
          size = .25) +  theme_bw() +
  theme(text = element_text(size = 15),
        axis.text = element_text(size = 6),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8)) +
  viridis::scale_fill_viridis(option = "inferno",
                              limits = c(0,20)) +
  labs(x = "", y = "", fill = "Mean\nabundance",
       title = "Scallop abundance (2000-2015)") + 
  theme(text = element_text(size = 14),
        title = element_text(size = 12))
)

# plot mean abundance, 1980-1999
(abund_per1 <- ggplot() +
  geom_tile(aes(x = lon, y = lat, fill = Years_1980_1999),
            # alpha = 1,
            data = change_plot,
            color = "black") +
  geom_sf(data = state_bounds, aes(), color = "gray",
          size = .25) +  theme_bw() +
  theme(text = element_text(size = 15),
        axis.text = element_text(size = 6),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8)) +
  viridis::scale_fill_viridis(option = "inferno",
                              limits = c(0,20)) +
  labs(x = "", y = "", fill = "Mean\nabundance",
       title = "Scallop abundance (1980-1999)") + 
  theme(text = element_text(size = 14),
        title = element_text(size = 12))
)

ggsave(plot = abund_per1,
       "figures/scallop_trawl_hd_mean_catch_1980_1999.png",
        height = 5, width = 5, dpi = 400)

ggsave(plot = abund_per2,
       "figures/scallop_trawl_hd_mean_catch_2000_2015.png",
        height = 5, width = 5, dpi = 400)

gg <- gridExtra::arrangeGrob(abund_per1, abund_per2, ncol = 2)

plot(gg)
ggsave(plot = gg, "figures/scallop_grid_mean_abund_2per.png",
        height = 5, width = 9, dpi = 400)
```
# Plot CHANGE in MEAN SBT by time period (0.5 grid cell)
```{r}
ggplot() +
  geom_tile(aes(x = lon, y = lat, fill = change),
            # alpha = 1,
            data = change_clim,
            color = "black") +
  geom_sf(data = state_bounds, aes(), color = "gray",
          size = .25) +  theme_bw() +
  theme(text = element_text(size = 15),
        axis.text = element_text(size = 6),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8)) +
  viridis::scale_color_viridis(option = "inferno") +
  labs(x = "", y = "", fill = "Absolute\nchange",
       title = "Change in sea bottom temp,\n(1980-1999 vs. 2000-2014)") + 
  scale_fill_gradient2(midpoint = 0, 
                       low = "steelblue", 
                       mid = "white",
                       high = "firebrick", space = "Lab") +
  theme(text = element_text(size = 14),
        title = element_text(size = 12))

# ggsave("figures/scallop_grid_change_in_mean_sbt.png",
#         height = 5, width = 5, dpi = 400)
```
# Plot CHANGE in SBT.actual (MORLEY) by time period (0.5 grid cell)
```{r}
ggplot() +
  geom_tile(aes(x = lon, y = lat, fill = change),
            # alpha = 1,
            data = change_morely,
            color = "black") +
  geom_sf(data = state_bounds, aes(), color = "gray",
          size = .25) +  theme_bw() +
  theme(text = element_text(size = 15),
        axis.text = element_text(size = 6),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8)) +
  # viridis::scale_color_viridis(option = "inferno") +
  labs(x = "", y = "", fill = "Absolute\nchange",
       title = "Change in sea bottom temp,\n(1980-1999 vs. 2000-2014)") + 
  scale_fill_gradient2(midpoint = 0, 
                       low = "steelblue", 
                       mid = "white",
                       high = "firebrick", space = "Lab") +
  theme(text = element_text(size = 14),
        title = element_text(size = 12))

ggsave("figures/scallop_grid_change_in_SBT.actual_MORLEY.png",
        height = 5, width = 5, dpi = 400)
```
# Plot CHANGE in MIN SBT by time period (0.5 grid cell)
```{r}
ggplot() +
  geom_tile(aes(x = lon, y = lat, fill = change),
            # alpha = 1,
            data = change_clim_min,
            color = "black") +
  geom_sf(data = state_bounds, aes(), color = "gray",
          size = .25) +  theme_bw() +
  theme(text = element_text(size = 15),
        axis.text = element_text(size = 6),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8)) +
  labs(x = "", y = "", fill = "Absolute\nchange",
       title = "Change in min sea bottom temp,\n(1980-1999 vs. 2000-2014)") + 
  scale_fill_gradient2(midpoint = 0, 
                       low = "steelblue", 
                       mid = "white",
                       high = "firebrick", space = "Lab") +
  theme(text = element_text(size = 14),
        title = element_text(size = 12))

# ggsave("figures/scallop_grid_change_in_min_sbt.png",
#         height = 5, width = 5, dpi = 400)
```
# Plot MEAN SBT by time period (0.5 grid cell)
```{r}
# plot mean temperature, 2000-2014
(per2 <- ggplot() +
  geom_tile(aes(x = lon, y = lat, fill = Years_2000_2015),
            # alpha = 1,
            data = change_clim,
            color = "black") +
  geom_sf(data = state_bounds, aes(), color = "gray",
          size = .25) +  theme_bw() +
  theme(text = element_text(size = 15),
        axis.text = element_text(size = 6),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8)) +
  viridis::scale_fill_viridis(option = "inferno") +
  labs(x = "", y = "", fill = "Mean\nbottom\ntemp (C)",
       size = "Mean CPUE\nin each\nperiod",
       title = "Mean sea bottom\ntemperature (2000-2014)") + 
  theme(text = element_text(size = 14),
        title = element_text(size = 12))
)

# plot mean temperature, 1980-1999
(per1 <- ggplot() +
  geom_tile(aes(x = lon, y = lat, fill = Years_1980_1999),
            # alpha = 1,
            data = change_clim,
            color = "black") +
  geom_sf(data = state_bounds, aes(), color = "gray",
          size = .25) +  theme_bw() +
  theme(text = element_text(size = 15),
        axis.text = element_text(size = 6),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8)) +
  viridis::scale_fill_viridis(option = "inferno") +
  labs(x = "", y = "", fill = "Mean\nbottom\ntemp (C)",
       size = "Mean CPUE\nin each\nperiod",
       title = "Mean sea bottom\ntemperature (1980-1999)") + 
  theme(text = element_text(size = 14),
        title = element_text(size = 12))
)

gg <- gridExtra::arrangeGrob(per1, per2, ncol = 2)

plot(gg)
ggsave(plot = gg, "figures/scallop_grid_mean_sbt_2per.png",
        height = 5, width = 9, dpi = 400)
```
# Plot CHANGE in MEAN O2 by time period (0.5 grid cell)
```{r}
# plot change in O2
ggplot() +
  geom_tile(aes(x = lon, y = lat, fill = change),
            # alpha = 1,
            data = change_O2,
            color = "black") +
  geom_sf(data = state_bounds, aes(), color = "gray",
          size = .25) +  theme_bw() +
  theme(text = element_text(size = 15),
        axis.text = element_text(size = 6),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8)) +
  labs(x = "", y = "", fill = "Absolute\nchange",
       title = "Change in oxygen,\n(1980-1999 vs. 2000-2014)") + 
  scale_fill_gradient2(midpoint = 0, 
                       low = "steelblue", 
                       mid = "white",
                       high = "firebrick", space = "Lab") +
  theme(text = element_text(size = 14),
        title = element_text(size = 12))

ggsave("figures/scallop_grid_change_in_mean_O2.png",
        height = 5, width = 5, dpi = 400)
```
# Plot MEAN O2 by time period (0.5 grid cell)
```{r}
# plot mean temperature, 2000-2014
(per2 <- ggplot() +
  geom_tile(aes(x = lon, y = lat, fill = Years_2000_2015),
            # alpha = 1,
            data = change_O2,
            color = "black") +
  geom_sf(data = state_bounds, aes(), color = "gray",
          size = .25) +  theme_bw() +
  theme(text = element_text(size = 15),
        axis.text = element_text(size = 6),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8)) +
  viridis::scale_fill_viridis(option = "inferno",
                              limits = c(0.00015, 0.00031)) +
  labs(x = "", y = "", fill = "Mean\noxygen",
       title = "Mean oxygen (2000-2014)") + 
  theme(text = element_text(size = 14),
        title = element_text(size = 12))
)

# plot mean temperature, 1980-1999
(per1 <- ggplot() +
  geom_tile(aes(x = lon, y = lat, fill = Years_1980_1999),
            # alpha = 1,
            data = change_O2,
            color = "black") +
  geom_sf(data = state_bounds, aes(), color = "gray",
          size = .25) +  theme_bw() +
  theme(text = element_text(size = 15),
        axis.text = element_text(size = 6),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8)) +
  viridis::scale_fill_viridis(option = "inferno",
                              limits = c(0.00015, 0.00031)) +
  labs(x = "", y = "", fill = "Mean\noxygen",
       title = "Mean oxygen (1980-1999)") + 
  theme(text = element_text(size = 14),
        title = element_text(size = 12))
)

gg <- gridExtra::arrangeGrob(per1, per2, ncol = 2)

ggsave(plot = gg, "figures/scallop_grid_mean_O2_2per.png",
        height = 5, width = 9, dpi = 400)
```
# General climate data plotting
```{r}
# Plot Morely climate data vs. new climate data
ggplot(clim.both) +
  geom_point(aes(x = SBT.min, y = min_temp)) +
  theme_bw() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  theme(text = element_text(size = 14))
ggsave("figures/newmin_vs_SBT.min.png")

ggplot(clim.both) +
  geom_point(aes(x = SBT.actual, y = mean_temp)) +
  theme_bw() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  theme(text = element_text(size = 14))
ggsave("figures/newmean_vs_SBT.actual.png")

ggplot(clim.both) +
  geom_point(aes(x = SBT.max, y = max_temp)) +
  theme_bw() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  theme(text = element_text(size = 14))
ggsave("figures/newmax_vs_SBT.max.png")

clim.both %>%
  filter(grid_lat > 39,
         grid_lat < 42,
         grid_lon < -70,
         grid_lon > -73) %>%
  ggplot() +
  geom_line(aes(x = year, y = min_temp, color = grid)) +
  geom_line(aes(x = year, y = SBT.min, color = grid),
            lty = 2) +
  facet_wrap(~grid_lat) +
  guides(color = "none") +
  theme_bw() +
  theme(text = element_text(size = 14)) +
  labs(y = "Minimum temperature", x = "")
ggsave("figures/newmin_vs_SBT.min_tsplot.png",
       height = 7, width = 7, dpi = 400)

clim %>%
  filter(grid %in% filter(change_clim)$grid) %>%
ggplot() + geom_line(aes(x = year, y = mean_temp, color = grid),
                     size = .5) + 
  guides(color = "none")
ggsave("figures/newmean_vs_year_tsplot.png")

clim %>%
  filter(grid %in% filter(change_clim)$grid) %>%
ggplot() + 
  # geom_line(aes(x = year, y = mean_temp, color = grid)) +
  geom_smooth(method = "lm", se = F, 
              aes(x = year, y = mean_temp, color = grid),
              size = .5) +
  guides(color = "none")
ggsave("figures/newmean_vs_year_trends.png")

ggplot(clim) + geom_line(aes(x = year, y = max_temp, color = grid)) + guides(color = "none")
ggsave("figures/newmax_vs_year_tsplot.png")

ggplot(clim) + geom_line(aes(x = year, y = min_temp, color = grid)) + guides(color = "none")
ggsave("figures/newmin_vs_year_tsplot.png")

ggplot(clim) + geom_line(aes(x = year, y = O2, color = grid)) + 
  guides(color = "none")
ggsave("figures/newO2_vs_year_tsplot.png")
```
# summarize mean depth by cell
```{r}
hd %>%
  group_by(grid, lat, lon) %>%
  summarize(mean_depth = mean(mean_depth),
            .groups = "drop") %>%
ggplot() +
  geom_tile(aes(x = lon, y = lat, fill = mean_depth), 
            alpha = 1, size = 0.25) +
  geom_sf(data = state_bounds, aes(), color = "gray",
          size = .25) +  theme_bw() +
  theme(text = element_text(size = 15),
        axis.text = element_text(size = 8)) +
  viridis::scale_fill_viridis(option = "inferno") +
  labs(x = "", y = "")

# ggsave("figures/scallop_trawl_hd_mean_depth.png",
#        height = 7, width = 7, dpi = 400)
```
# summarize mean bottom temperature by cell
Trawl survey data
```{r}
hd %>%
  group_by(grid, lat, lon) %>%
  summarize(mean_bottemp = mean(mean_bottemp, na.rm = T),
            .groups = "drop") %>%
ggplot() +
  geom_tile(aes(x = lon, y = lat, fill = mean_bottemp), 
            alpha = 1, size = 0.25) +
  geom_sf(data = state_bounds, aes(), color = "gray",
          size = .25) +  theme_bw() +
  theme(text = element_text(size = 15),
        axis.text = element_text(size = 8)) +
  viridis::scale_fill_viridis(option = "inferno") +
  labs(x = "", y = "")

# ggsave("figures/scallop_trawl_hd_mean_bottemp.png",
#        height = 7, width = 7, dpi = 400)
```
# plot time series for cells
```{r}
ts_plot_data <- hd %>%
  group_by(grid) %>%
  mutate(mean_hauls = mean(n_hauls),
         mean_bt = mean(mean_bottemp, na.rm = T),
         n_years = length(n_hauls),
         frac_zeros = sum(mean_wtcpue==0)/n_years) %>%
  filter(mean_hauls >= 1,
         n_years > 10,
         n_hauls > 0,
         frac_zeros < 0.75) %>%
  st_drop_geometry() %>%
  as.data.frame() %>%
  group_by(grid, lat, mean_bt, year, decade) %>%
  summarize(mean_wtcpue = mean(mean_wtcpue),
            mm_0_70 = mean(mean_0_70),
            mm_70_100 = mean(mean_70_100),
            mm_100_up = mean(mean_100_up),
            .groups = "drop") %>%
  pivot_longer(cols = mm_0_70:mm_100_up,
               names_to = "length",
               values_to = "num") %>%
  mutate(lat_band = fct_reorder(as.factor(trunc(lat)), 
                                as.numeric(trunc(lat)),
                                .desc = T),
         order = case_when(length == "mm_0_70" ~ 1,
                           length == "mm_70_100" ~ 2,
                           length == "mm_100_up" ~ 3),
         length = fct_reorder(length, order))

ts_plot_data  %>%
  ggplot() +
  geom_line(aes(x = as.numeric(year), y = log10(num+1), 
                color = mean_bt, group = grid)) +
  facet_grid(rows = vars(lat_band), cols = vars(length)) +
  viridis::scale_color_viridis() +
  labs(x = "") +
  theme_bw() +
  theme(text = element_text(size = 13)) +
  scale_x_continuous(breaks = c(1980, 2000, 2019),
                     labels = c(1980, 2000, 2019))

# ggsave("figures/scallop_trawl_hd_mean_catch_by_year.png",
#        height = 7, width = 10, dpi = 400)
```
# Plot 0.5 x 0.5 degree averages - mean catch by decade & size
```{r}
# summarize mean catch by decade and size
readRDS("data_noaa/trawl/survdat.rds") %>%
  # filter to just scallops (401)
  filter(SVSPP == 401) %>%
  # divide into length bins (measures coarser than dredge survey)
  mutate(
    n_0_40 = ifelse(LENGTH <= 4, NUMLEN, 0),
    n_40_80 = ifelse(LENGTH > 4 & LENGTH < 8, NUMLEN, 0),
    n_80_up = ifelse(LENGTH >= 8, NUMLEN, 0),
    station_temp = case_when(nchar(STATION)==1 ~ 
                               paste0("00", as.character(STATION)),
                             nchar(STATION)==2 ~
                               paste0("0", as.character(STATION)),
                             nchar(STATION)==3 ~
                               as.character(STATION),
                             TRUE ~ "problem"),
    haulid = paste(CRUISE6,station_temp,STRATUM, sep = "-")) %>%
  select(-station_temp) %>%
# include total catch number from tlen in hauls_df
  group_by(haulid) %>%
  summarize(n_catch_all = sum(NUMLEN, na.rm = T), # 23 NA, 1963: 17
            n_0_40 = sum(n_0_40),
            n_40_80 = sum(n_40_80), 
            n_80_up = sum(n_80_up), 
            .groups = "drop") %>%
  right_join(hauls_df_step1, by = "haulid") %>%
  replace_na(replace = list(n_catch_all = 0,
                            n_0_40 = 0,
                            n_40_80 = 0,
                            n_80_up = 0)) %>%
  arrange(haulid, year) %>%
#note: haulid "201404-211-1100" has EXTREMELY high count (>150k)
# summarizing hauls data by 0.5 degree grid cell
    mutate(lat = grid_lat,
         lon = grid_lon) %>%
    group_by(grid, year, decade, lat, lon, 
           grid_lat, grid_lon) %>%
  summarize(mean_wtcpue = mean(wtcpue),
            mean_tot_num = mean(n_catch_all),
            mean_0_40 = mean(n_0_40),
            mean_40_80 = mean(n_40_80),
            mean_80_up = mean(n_80_up),
            n_hauls = length(wtcpue),
            mean_depth = mean(elev, na.rm = T),
            mean_bottemp = mean(bottemp, na.rm = T),
            .groups = "drop") %>%
  arrange(grid, year) %>%
  filter(lat >34) %>%
# make data frames into sf objects and load spatial data
  st_as_sf(coords = c("grid_lon", "grid_lat"),
           crs = 4326) %>%
  group_by(grid) %>%
  mutate(mean_hauls = mean(n_hauls),
         mean_bt = mean(mean_bottemp, na.rm = T),
         n_years = length(n_hauls),
         frac_zeros = sum(mean_wtcpue==0)/n_years) %>%
  st_drop_geometry() %>%
  as.data.frame() %>%
  group_by(grid, lat, lon, mean_bt, year, decade) %>%
  summarize(mean_wtcpue = mean(mean_wtcpue),
            mm_0_40 = mean(mean_0_40),
            mm_40_80 = mean(mean_40_80),
            mm_80_up = mean(mean_80_up),
            .groups = "drop") %>%
  rename("Size 0-40 mm" = mm_0_40,
         "Size 40-80 mm" = mm_40_80,
         "Size >80 mm" = mm_80_up) %>%
  pivot_longer(cols = starts_with("Size"),
               names_to = "length",
               values_to = "num") %>%
  mutate(order = case_when(length == "Size 0-40 mm" ~ 1,
                           length == "Size 40-80 mm" ~ 2,
                           length == "size >80 mm" ~ 3)) %>%
  mutate(length = forcats::fct_reorder(length, order)) %>%
ggplot() +
  geom_tile(aes(x = lon, y = lat, fill = log10(num+1)), 
            alpha = 1, size = 0.25) +
  geom_sf(data = state_bounds, aes(), color = "gray",
          size = .25) +  theme_bw() +
  theme(text = element_text(size = 15),
        axis.text = element_text(size = 6),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8)) +
  viridis::scale_fill_viridis(option = "inferno") +
  labs(x = "", y = "", fill = "log10(num+1)") +
  facet_grid(rows = vars(decade), cols = vars(length))

# ggsave("figures/scallop_trawl_hd_mean_catch_by_decade_length.png",
#         height = 7, width = 7, dpi = 400)
```